Error Reporting Parsers: a Monad Transformer Approach

### 2013-05-18 article.tex
improve article section introducing parsers

\subsection{Parser-specific combinators}
\begin{verbatim}
item :: Parser t t
item = Parser f
  where 
    f [] = Nothing
    f (x:xs) = Just (xs, x)

literal :: Eq t => t -> Parser t t
literal x = afilter (== x) item
\end{verbatim}
Next is the \verb+item+ parser, the simplest parser we'll deal with.  All it does is 
consume a single token if available, and fail otherwise.

@@

For choosing between alternatives, the Alternative type class will do, so we 
create an instance that is defined in terms of the Alternative instance of the
underlying result type, \verb+Maybe+.
\begin{verbatim}
instance Alternative (Parser t) where
  empty = Parser (const empty)
  Parser l <|> Parser r = Parser (\ts -> l ts <|> r ts)

-- similar to mfilter, but for Alternative instead of MonadPlus
check :: (Monad f, Alternative f) => (a -> Bool) -> f a -> f a
check p m =
    m >>= \x -> if p x then return x else empty
\end{verbatim}

>>>>

@@

We use the Alternative type class for expressing choice; the Parser instance is  
is defined in terms of the Alternative instance of the
underlying result type, \verb+Maybe+.
\begin{verbatim}
instance Alternative (Parser t) where
  empty = Parser (const empty)
  Parser l <|> Parser r = Parser (\ts -> l ts <|> r ts)
\end{verbatim}

\subsection{Parser-specific combinators}
\verb+item+ is a very simple parser; it 
consumes a single token if available, and fails otherwise:
\begin{verbatim}
item :: Parser t t
item = Parser f
  where 
    f [] = Nothing
    f (x:xs) = Just (xs, x)
\end{verbatim}

Analagous to \verb+mfilter+ is the \verb+check+ combinator, which is used to
validate a parse result:
\begin{verbatim}
check :: (Monad f, Alternative f) => (a -> Bool) -> f a -> f a
check p m =
    m >>= \x -> if p x then return x else empty
\end{verbatim}

Using \verb+check+, we can build a very convenient combinator to parse specific
matching tokens: 
\begin{verbatim}
literal :: Eq t => t -> Parser t t
literal x = check (== x) item
\end{verbatim}
